Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-07-27 Moltres.tex
Correct precursor scaling.
Begin MSRE comparison.

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{0.8}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\textfraction}{0.07}
\renewcommand{\floatpagefraction}{0.7}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}
% hyperref usually has to go last
\usepackage[hidelinks]{hyperref}
% but glossaries behaves best if after hyperref
\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
% cleveref only behaves if after hyperref & glossaries
\usepackage{cleveref}
\usepackage{minted}
\usepackage{ifthen}
%% \usepackage{listings}  % Allows code printing
%% \definecolor{codegreen}{rgb}{0,0.6,0}
%% \definecolor{codegray}{rgb}{0.5,0.5,0.5}
%% \definecolor{codepurple}{rgb}{0.58,0,0.82}
%% \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
%% \lstdefinestyle{mystyle}{
%%     backgroundcolor=\color{backcolour},
%%     commentstyle=\color{codegreen},
%%     keywordstyle=\color{magenta},
%%     numberstyle=\tiny\color{codegray},
%%     stringstyle=\color{codepurple},
%%     basicstyle=\footnotesize,
%%     breakatwhitespace=false,
%%     breaklines=true,
%%     captionpos=b,
%%     keepspaces=true,
%%     numbers=left,
%%     numbersep=5pt,
%%     showspaces=false,
%%     showstringspaces=false,
%%     showtabs=false,
%%     tabsize=2
%% }
%% \lstset{style=mystyle}

@@

Building on the massively parallelizable \gls{MOOSE} framework allows Moltres
to be run on super-computing platforms like \gls{NCSA}'s Blue Waters. For some
three-dimensional simulations (results not discussed here), the number of
elements in the mesh and total number of degrees of freedom exceed one million
and ten million respectively. To handle problems of this size, Moltres has been
run on up to 608 cores. However, reducing the problem dimension from three to
two and using a structured mesh that can be much more coarse in the axial
direction allows problem solution to be carried out on a single core in under
five minutes.

@@

As a proof of concept and to ease resolution of strong transient gradients in
temperature between fuel and moderator, the physical three dimensional problem
is reduced to a two dimensional axisymmetric model. The model molten salt
reactor is based on the \gls{MSRE}. To approximately simulate the lattice
structure of the \gls{MSRE}, a geometry is constructed with 14 repeating
fuel-moderator regions, as shown in \Cref{fig:geom}. The fuel and moderator
radii are chosen such that the resulting area/volume fraction of fuel is .225 as
for the \gls{MSRE}.

@@

\Cref{fig:flux_comparison} compares the fast and thermal neutron fluxes at the
reactor mid-plane ($z=H/2$). Local thermal flux growth and fast flux decay in
moderator regions and visa versa in fuel regions are even more apparent. The
larger fast flux magnitude is attributable to using \gls{BOL} fuel
composition. As the fuel salt reaches an equilibrium composition, the thermal
flux is expected to exceed that of the fast group.

@@

% Fix me! Precursor concentrations have goofy scaling
\begin{figure}[htpb]
  \centering
  \includegraphics{2d_gamma_heating_pre1.eps}
        \caption{The concentration of the group of longest lived precursors
        ($\lambda = 1.24\cdot 10^{-2} s^{-1}$) peaks near the reactor outlet
        in this 2-D axisymmetric model ($r=0$ is center of core).}
  \label{fig:pre1}
\end{figure}

@@

% Fix me! Precursor concentrations have goofy scaling
\begin{figure}[htpb]
  \centering
  \includegraphics{2d_gamma_heating_pre6.eps}
        \caption{The concentration of the group of shortest lived precursors
        ($\lambda = 3.07 s^{-1}$) peaks near the reactor center
        in this 2-D axisymmetric model ($r=0$ is center of core).}
  \label{fig:pre6}
\end{figure}

\subsection{3D}

@@

@@

\begin{figure}[htpb]
  \centering
  \includegraphics{fuel_vs_moderator_axial_sweep.eps}
        \caption{Moltres predicted axial temperature profiles near r = 0}
  \label{fig:moltres_axial_temps}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics{msre_axial_temps.png}
        \caption{\gls{MSRE} design axial temperature profiles near r = 0}
  \label{fig:msre_axial_temps}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{fast_vs_thermal_nt_flux_msre.png}
  \caption{\gls{MSRE} design radial temperature profiles at reactor mid-plane}
  \label{fig:msre_radial_temps}
\end{figure}

>>>>

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{0.8}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\textfraction}{0.07}
\renewcommand{\floatpagefraction}{0.7}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}
% hyperref usually has to go last
\usepackage[hidelinks]{hyperref}
% but glossaries behaves best if after hyperref
\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
% cleveref only behaves if after hyperref & glossaries
\usepackage{cleveref}
\usepackage{minted}
\usepackage{ifthen}
\usepackage{subcaption}
%% \usepackage{listings}  % Allows code printing
%% \definecolor{codegreen}{rgb}{0,0.6,0}
%% \definecolor{codegray}{rgb}{0.5,0.5,0.5}
%% \definecolor{codepurple}{rgb}{0.58,0,0.82}
%% \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
%% \lstdefinestyle{mystyle}{
%%     backgroundcolor=\color{backcolour},
%%     commentstyle=\color{codegreen},
%%     keywordstyle=\color{magenta},
%%     numberstyle=\tiny\color{codegray},
%%     stringstyle=\color{codepurple},
%%     basicstyle=\footnotesize,
%%     breakatwhitespace=false,
%%     breaklines=true,
%%     captionpos=b,
%%     keepspaces=true,
%%     numbers=left,
%%     numbersep=5pt,
%%     showspaces=false,
%%     showstringspaces=false,
%%     showtabs=false,
%%     tabsize=2
%% }
%% \lstset{style=mystyle}

@@

Building on the massively parallelizable \gls{MOOSE} framework allows Moltres
to be run on super-computing platforms like \gls{NCSA}'s Blue Waters. For some
three-dimensional simulations, the number of
elements in the mesh and total number of degrees of freedom exceed one million
and ten million respectively. To handle problems of this size, Moltres has been
run on up to 608 cores. However, reducing the problem dimension from three to
two and using a structured mesh that can be much more coarse in the axial
direction allows problem solution to be carried out on a single core in under
five minutes.

@@

The model molten salt
reactor is based on the \gls{MSRE}. When developing new physics or investigating
different types of transients, the model problem can be reduced to a two
dimensional axisymmetric one for rapid proof of concept. To approximately simulate the lattice
structure of the \gls{MSRE} under 2D conditions, a geometry is constructed with 14 repeating
fuel-moderator regions, as shown in \Cref{fig:geom}. The fuel and moderator
radii are chosen such that the resulting area/volume fraction of fuel is .225 as
for the \gls{MSRE}.

@@

\Cref{fig:flux_comparison} compares the fast and thermal neutron fluxes at the
reactor mid-plane ($z=H/2$). Local thermal flux growth and fast flux decay in
moderator regions and visa versa in fuel regions are even more apparent. The
larger fast flux magnitude is attributable to using \gls{BOL} fuel
composition. As the fuel salt reaches an equilibrium composition, the fast to
thermal flux ratio is expected to decrease.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics{2d_gamma_heating_pre1_scaled.eps}
        \caption{The concentration of the group of longest lived precursors
        ($\lambda = 1.24\cdot 10^{-2} s^{-1}$) peaks near the reactor outlet
        in this 2-D axisymmetric model ($r=0$ is center of core).}
  \label{fig:pre1}
\end{figure}

@@

\begin{figure}[htpb]
  \centering
  \includegraphics{2d_gamma_heating_pre6_scaled.eps}
        \caption{The concentration of the group of shortest lived precursors
        ($\lambda = 3.07 s^{-1}$) peaks near the reactor center
        in this 2-D axisymmetric model ($r=0$ is center of core).}
  \label{fig:pre6}
\end{figure}

\subsection{Three dimensional simulation capability}

\Cref{fig:3d_group1,fig:3d_temp} show Moltres physics applied to a three
dimensional geometry. The fast group flux (\cref{fig:3d_group1}) is in good
qualitative agreement with the two dimensional axisymmetric case shown in
\cref{fig:group1}. \Cref{fig:3d_temp} is in similarly good agreement with
\cref{fig:temp}. \Cref{fig:3d_temp_fuel_outlet} shows the temperature profile at
the outlet of the reactor (z = H). This three dimensional case contained
1,155,045 degrees of freedom and took only 2.5 hours to solve on 160 Blue Waters
cores.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics{3d_gamma_heating_z_slice_temp.eps}
        \caption{Temperature at the reactor outlet for 3D \gls{MSRE}-like
          model.}
  \label{fig:3d_temp_fuel_outlet}
\end{figure}

@@

\Cref{fig:temp_compare} shows a comparison between Moltres predicted temperature
profiles with cosinusoidal gamma heating and \gls{MSRE} design models
\cite{briggs_molten-salt_1964} along the reactor center-line (r = 0).

\begin{figure}[htpb]
    \centering
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{fuel_vs_moderator_axial_sweep.eps}
      \caption{Moltres predicted axial temperature profiles near r = 0}
      \label{fig:moltres_axial_temps}
    \end{subfigure}
    %\quad
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_temps.png}
      \caption{\gls{MSRE} design axial temperature profiles near r = 0
        \cite{briggs_molten-salt_1964}}
      \label{fig:msre_axial_temps}
    \end{subfigure}
    \caption{Comparison of \gls{MSRE} design and Moltres predicted axial
      temperature profiles}
    \label{fig:temp_compare}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{msre_radial_fluxes.png}
  \caption{\gls{MSRE} design radial flux profiles at reactor mid-plane
    \cite{briggs_molten-salt_1964}}
  \label{fig:msre_radial_fluxes}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_fluxes.png}
  \caption{\gls{MSRE} design axial flux profiles at r = 0
    \cite{briggs_molten-salt_1964}}
  \label{fig:msre_axial_fluxes}
\end{figure}
